clusters

### 2015-07-08 cluster.rmd
more clust

library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')

# library('knitr')
# library('ggplot2')
# library("dplyr")
# library("plyr")
# library('magrittr')
# library('reshape2')
# library("rpart")
# library('car')
# library('e1071')
# library('corrgram')
# library('party')
# library('dendextend')
# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")

@@

code.contact
code.client
Project.Value
JD.Second
Billing.Type
Post.Code
code.ProjEng
Billing.Type
majority.pos

@@

#Q-Q plot of each variable
clust<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, pc.pro, client.totinv,
                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
                         cost.mlsto, pc.majpos, return.pdol)

clust[clust$timespan ==0,]$timespan <- 1
clust[clust$pc.pro ==0,]$pc.pro <- 5
clust[clust$hours.perday ==0,]$hours.perday <- .1
clust[clust$hours.perday ==0,]$hours.perday <- .1
clust[clust$balance.mlsto ==0,]$balance.mlsto <- 1
clust[clust$hrs.mlsto ==0,]$hrs.mlsto <- 1
clust[clust$cost.mlsto ==0,]$cost.mlsto <- 1
clust[clust$return.pdol ==0,]$return.pdol <- .01

for(i in 1:ncol(clust)){
qqPlot(clust[,i],
       main=colnames(clust)[i])
}

#qqplots for log of each variable
for(i in 2:ncol(clust)+1){
qqPlot(clust[,i] %>% log,
       main=colnames(clust)[i])
}

#qqplots for sqrt of each variable
for(i in 2:ncol(clust[,-1])+1){
qqPlot(clust[,i] %>% sqrt,
       main=colnames(clust)[i])
}

#density plots - plain variables
for(i in 1:ncol(clust)){
print(
        ggplot(clust, aes_string(x=colnames(clust[i]))) + geom_histogram(aes(y=..density..))
        )
}
#density plots - log transformed variables
clust.log<- log(clust)
for(i in 1:ncol(clust.log)){
        print(
                ggplot(clust.log, aes_string(x=colnames(clust.log[i]))) + geom_histogram(aes(y=..density..))
                )
}

#plot sqrt of timespan
clust$timespan <- clust$timespan %>% sqrt
ggplot(clust, aes_string(x='timespan')) + geom_histogram(aes(y=..density..))

@@

clust.summary[,c(1,12,10,11, 13, 9,2,7,6,3,4,8,5,14:16)]

@@

>>>>

library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')

# library('knitr')
# library('ggplot2')
# library("dplyr")
# library("plyr")
# library('magrittr')
# library('reshape2')
# library("rpart")
# library('car')
# library('e1071')
# library('corrgram')
# library('party')
# library('dendextend')
# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
# opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")

@@

code.contact
code.client
Project.Value
JD.Second
Billing.Type
Post.Code
code.ProjEng
Billing.Type
majority.pos
pc.majpos

@@

# #Q-Q plot of each variable
# clust<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, pc.pro, client.totinv,
#                          Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
#                          cost.mlsto, pc.majpos, return.pdol)
# 
# #get rid of zeroes for log transform
# 
# clust[clust$timespan ==0,]$timespan <- 1
# clust[clust$pc.pro ==0,]$pc.pro <- 5
# clust[clust$hours.perday ==0,]$hours.perday <- .1
# clust[clust$hours.perday ==0,]$hours.perday <- .1
# clust[clust$balance.mlsto ==0,]$balance.mlsto <- 1
# clust[clust$hrs.mlsto ==0,]$hrs.mlsto <- 1
# clust[clust$cost.mlsto ==0,]$cost.mlsto <- 1
# clust[clust$return.pdol ==0,]$return.pdol <- .01
# 
# for(i in 1:ncol(clust)){
# qqPlot(clust[,i],
#        main=colnames(clust)[i])
# }
# 
# #qqplots for log of each variable
# for(i in 2:ncol(clust)+1){
# qqPlot(clust[,i] %>% log,
#        main=colnames(clust)[i])
# }
# 
# #qqplots for sqrt of each variable
# for(i in 2:ncol(clust[,-1])+1){
# qqPlot(clust[,i] %>% sqrt,
#        main=colnames(clust)[i])
# }
# 
# #density plots - plain variables
# for(i in 1:ncol(clust)){
# print(
#         ggplot(clust, aes_string(x=colnames(clust[i]))) + geom_histogram(aes(y=..density..))
#         )
# }
# #density plots - log transformed variables
# clust.log<- log(clust)
# for(i in 1:ncol(clust.log)){
#         print(
#                 ggplot(clust.log, aes_string(x=colnames(clust.log[i]))) + geom_histogram(aes(y=..density..))
#                 )
# }
# 
# #plot sqrt of timespan
# clust$timespan <- clust$timespan %>% sqrt
# ggplot(clust, aes_string(x='timespan')) + geom_histogram(aes(y=..density..))

@@

clust.summary<- clust.summary[,c(1,12,10,11, 13, 9,2,7,6,3,4,8,5,14:16)]

# add medians
all7c.med<- ddply(kmeans.plot, .(groups), numcolwise(median))

all7c.med<- all7c.med[, c(2,13,11,12,14,10,3,8,7,4,5,9,6)]

@@

Very interesting, now want to optimise number of clusters- refer wards tutorial

```{r optimise clustnum,echo=FALSE}
library('fpc', lib='C:/Progra~1/R/R-3.2.0/library')
library('vegan', lib = 'C:/Progra~1/R/R-3.2.0/library')

#try clustering a low number of numeric variables, trialling iffy ones to see which ones are good for clustering
clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
                                      hrs.mlsto), 2, 20, iter= 100)
plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
# client.totinv indicating about 4 clusters
# Num.days - good, indicating 18 custers
# hrs.mlsto - ok, heavy dip, good at about 17 clusters
# pc.pro - 5 groups good, still iffy
# mean.peeps - good at about 4 clusters
# Num.disc - ok, inidicating 12 groups
# timespan: not good for clustering
# no.users: not good for clustering
# hours.perday: no good for clustering

## try out all iffy variables together, eliminate duds
clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
                                      client.totinv, Num.days, hrs.mlsto, pc.pro, mean.peeps,
                                      Num.disc), 2, 20, iter= 100)
plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
# all together, no good
# smaller subset of variables I am most interested in that still works for clustering:

clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
                                      hrs.mlsto
                                      ), 2, 20, iter= 100)
plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)

# 17 clusters
# Cluster cohesion: create plots of SSE index (distance within clusters - sum of squared error) 
# then calinksi (index of cluster homogeneity and distance where higher number is better)
# calinski: between cluster distance/within cluster distance
ind.plot<- melt(clust.km$results)
ggplot(ind.plot %>% filter(Var1 == 'SSE'), aes(x=Var2, y=value, group=1)) + 
        geom_line( ) +
        theme(axis.text.x=element_text(angle=45,hjust=1))
ggplot(ind.plot %>% filter(Var1 == 'calinski'), aes(x=Var2, y=value, group=1)) + 
        geom_line( )+
        theme(axis.text.x=element_text(angle=45,hjust=1))
# 17 clusters should be good

Analysis shows that inv.mlsto, balance.mlsto, return.pdol, and hrs.mlsto should be used for clustering and that 17 clusters is ideal.
I'm not completely happy with the clustering.
I think I should now add a new variable: inv.mlsto x return.pdol
This will give a 'success' scale, weighing large jobs even with a smaller return.pdol as quite good and tiny jobs with a good return.pdol as average to low.

this means I will need to check normality, scale the variable before adding it to clust

```{r new var, echo=FALSE}
#new variable for all7c - success
# success= return.pdolxinv.mlsto

all7c<- transform(all7c, suc= inv.mlsto*return.pdol)
all7c$suc %>% summary

all7c %>% arrange(-suc) %>% head

qqPlot(all7c$suc,
       main='success')

qqPlot((all7c$suc+102700) %>% log)

#can't log because of negatives

#density plots - plain variables

ggplot(all7c, aes_string(x='suc')) + geom_histogram(aes(y=..density..))

#density plots - log transformed variables

trial<- (all7c$suc) %>% as.data.frame
ggplot(trial, aes_string(x='.')) + geom_histogram(aes(y=..density..))

### CAN WE TAKE LOG OF ABSOLUTE VALUES AND THEN MAKE NUMBERS THAT WERE PREVIOUSLY NEG NEG AGAIN?

clust$suc<- scale(all7c$suc)

Repeat clustering with new suc

```{r new successvar, echo=FALSE}

#investigate new variable a little

clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
                                      suc, pc.pro, hrs.mlsto
                                      ), 2, 20, iter= 100)
plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)

# 10 clusters
# Cluster cohesion: create plots of SSE index (distance within clusters - sum of squared error) 
# then calinksi (index of cluster homogeneity and distance where higher number is better)
# calinski: between cluster distance/within cluster distance
ind.plot<- melt(clust2.km$results)
ggplot(ind.plot %>% filter(Var1 == 'SSE'), aes(x=Var2, y=value, group=1)) + 
        geom_line( ) +
        theme(axis.text.x=element_text(angle=45,hjust=1))
ggplot(ind.plot %>% filter(Var1 == 'calinski'), aes(x=Var2, y=value, group=1)) + 
        geom_line( )+
        theme(axis.text.x=element_text(angle=45,hjust=1))
# 17 clusters should be good
```
